This page and the next page should be removed

### 2022-01-20 analysis/manuscript.Rmd
Initialize new pipeline analysis

---
title             : |
  | Event-Related Potentials of the Semantically Informed
  | Perception of Unfamiliar Objects
shorttitle        : "Semantic Knowledge and Unfamiliar Objects"

author: 
  - name          : "Alexander Enge"
    affiliation   : "1,2"
  - name          : "Franziska Süß"
    affiliation   : "3"
  - name          : "Rasha Abdel Rahman"
    affiliation   : "1"
    corresponding : yes
    address       : "Rudower Chaussee 18, 12489 Berlin"
    email         : "rasha.abdel.rahman@hu-berlin.de"

affiliation:
  - id            : "1"
    institution   : "Humboldt-Universität zu Berlin"
  - id            : "2"
    institution   : "Max Planck Institute for Human Cognitive and Brain Sciences"
  - id            : "3"
    institution   : "Fachhochschule des Mittelstands"

authornote: |
  \addORCIDlink{Alexander Enge}{0000-0003-0100-2297}

@@

  The preprocessed data and code for this study are openly available at https://osf.io/uksbc/.

  We have no conflict of interest to disclose.

abstract: |
  Does our perception of an object change as soon as we discover what function it serves? This question is relevant not only for our everyday lives, where we may encounter novel tools and gadgets as parts of our dynamic working and private environments; it also pertains to the long-standing debate around the (im)penetrability of perception by higher cognitive capacities. In this experiment, we showed participants (*n* = 24) pictures of 120 unfamiliar objects either together with matching information about their function---leading to semantically informed perception---or together with non-matching information---resulting in naive perception. We measured event-related potentials (ERPs) to investigate at which stages in the visual processing hierarchy these two types of object perception differed from one another. We found that semantically informed as compared to naive perception was associated with larger amplitudes in the N170 component and reduced amplitudes in the N400 component. When the same objects were presented once more (without any information), the N400 effect persisted and we now also observed enlarged amplitudes in the P1 component in response to objects for which semantically informed perception had taken place. We replicated these novel findings in an independent sample (*n* = 24). Consistent with previous work, these results suggest that obtaining semantic information about previously unfamiliar objects alters aspects of their lower-level visual perception (P1 component), higher-level visual perception (e.g., holistic perception; N170 component), and semantic processing (N400 component).

keywords          : "objects, semantic knowledge, visual perception, event-related potentials"

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

bibliography      : "manuscript_files/r-references.bib"
csl               : "manuscript_files/apa.csl"
zotero            : "aha"
documentclass     : "apa7"
classoption       : "man"
output            :
  papaja::apa6_pdf:
    latex_engine  : "xelatex"

editor_options: 
  chunk_output_type: console

header-includes:
  - \geometry{a4paper}
  - \fancyheadoffset[R,L]{0pt}
  - \raggedbottom
  - \usepackage[all]{nowidow}
  - \usepackage[bottom]{footmisc}
  - \interfootnotelinepenalty=10000
  - \usepackage{setspace}
  - \AtBeginEnvironment{tabular}{\doublespacing}
  - \usepackage{makecell}
  - \renewcommand{\cellset}{\renewcommand{\arraystretch}{0.7}}
  - \captionsetup[table]{font={stretch=1.5}}
  - \captionsetup[figure]{font={stretch=1.5}}
---

```{r setup, include=FALSE}
# Do you want to re-run the EEG preprocessing? Note that this can only be done with the raw data,
# which are available from the corresponding author (rasha.abdel.rahman@hu-berlin.de) upon request.
rerun_preproc <- TRUE

# Do you want to re-run the linear mixed-effect models? This can be done even if only the
# preprocessed single-trial data are available (from GitHub and/or the OSF). Note, however, that
# fitting all models may take 24 hours or more.
rerun_models <- FALSE

# Do you want to run the analysis for Experiment 3?
run_exp3 <- FALSE

# Load packages
library(papaja)
library(reticulate)
library(tidyverse)
library(magrittr)
library(cowplot)

# Set global chunk options
knitr::opts_chunk$set(include = FALSE, fig.align = "center", out.width = "100%", fig.width = 10)

# Link to the bibliography file
r_refs("analysis/manuscript_files/r-references.bib")

# Source plotting functions
source("analysis/manuscript_files/functions_plots_tables.R")

# Create an export folder for results
dir.create("analysis/export", showWarnings = FALSE)

# Define ERP components of interest
comps <- tibble(
  name = c("P1", "N170", "N400"),
  tmin = c(100, 150, 400),
  tmax = c(150, 200, 700),
  roi = list(
    c("PO3", "PO4", "POz", "O1", "O2", "Oz"),
    c("P7", "P8", "PO7", "PO8", "PO9", "PO10"),
    c("C1", "C2", "Cz", "CP1", "CP2", "CPz")
  )
)

# Define some filenames for I/O
fnames <- list(
  epochs = "analysis/export/aha-epo.fif",
  trials = "analysis/export/aha-trials.csv",
  gavgs = "analysis/export/aha-gavgs.RDS",
  conditions = "analysis/export/aha-obj-per-cond.csv",
  montage = "analysis/export/aha-montage.csv"
)
```

Does our perception of an object change as soon as we discover what function it serves? In this study, we presented participants with unfamiliar objects that were preceded either by matching information about their function, leading to semantically informed perception, or by non-matching information, leading to naive perception. We expected semantically informed perception to differ from naive perception within the object processing hierarchy and we capitalized on the high temporal resolution of event-related potentials (ERPs) to find out at which stage (or stages) this was indeed the case. This allowed us to show for the first time that knowledge about the function of a previously unfamiliar object alters its early cortical processing in the same trial as it is being discovered and within less than 200 ms after the presentation of the object.

The claim that higher-level cognitive capacities such as knowledge or language can modulate lower-level perceptual processes has been the source of a long-running debate. This debate revolves around the question of whether or not our perception of the (visual) world is "cognitively impenetrable" [@pylyshyn1999] or, in other words, determined solely by the visual input reaching our retina. According to one view, higher-level mental states and functions are not able to penetrate perception and kick in only after lower-level perceptual analysis of the visual input has taken pace [e.g., @firestone2016; @fodor1983; @pylyshyn1999]. Visual perception itself is treated as an encapsulated module that processes the retinal input in a feed-forward fashion, progressing from lower areas with smaller receptive field sizes to areas representing increasingly complex shapes and, eventually, whole objects [@dicarlo2012; @marr1982].

This position is challenged by the alternative view that perceptual processing dynamically interacts with other aspects of cognition from early on [e.g., @churchland1994; @lupyan2015; @teufel2017]. The myriad feedback connections from areas higher up the visual hierarchy (e.g., the middle temporal and lateral occipital cortices) to early visual areas [e.g., V1 and V2\; @bullier2001; @gilbert2013]---as well as behavioral evidence---have inspired theories of visual perception that emphasize the top-down influence of cognitive processes. For instance, the reverse hierarchy theory [@ahissar2004; @hochstein2002] assumes that conscious visual perception occurs initially at the level of whole objects or object categories. Only after this high-level interpretation of the stimulus has been obtained, its more for fine grained visual details---if relevant for the current task---are being accessed from lower-level areas via top-down connections. Along similar lines, the label feedback hypothesis [@lupyan2012] posits that the activation of an object's name (i.e., a high level property) transiently warps perceptual space so that its diagnostic visual features are being processed preferentially. Finally, an active role of top-down influences is also promoted by theories of vision in the framework of Bayesian inference and predictive coding [e.g., @clark2013; @lupyan2015; @panichello2013; @yuille2006].

Empirical findings from experimental psychology have recently added additional support to the view that perception can indeed be penetrated by other aspects of cognition [for review, see @collins2014; @vetter2014]. These aspects of cognition may include transient states such as emotions [e.g., @bocanegra2009; @phelps2016] or intentions [e.g., @balcetis2010; @cole2012] as well as learned capacities such as language [e.g., @boutonnet2015; @maier2018; @mo2011] or declarative knowledge about faces [e.g., @abdelrahman2011; @suess2015; @eiserbeck2020] and objects [e.g., @abdelrahman2008; @gauthier2003; @weller2019]. Despite this wealth of empirical evidence, critics of the idea that cognitive functions can penetrate perception have remained skeptical and pointed out important methodological shortcomings in large parts of this literature [@firestone2016; @machery2015]. They argue, for example, that most behavioral paradigms have not been able to discern between perceptual and post-perceptual (e.g., memory-related) effects. Furthermore, cognitive influences in some studies were confounded with differences in the low-level visual input itself. We will revisit these and other potential pitfalls with regards to the present study in the Discussion section once we have laid out our methodology and results.

In general, however, it is important to note that one productive way to circumvent at least some of these concerns is to use neurophysiological recording methods such as ERPs [@luck2014]: Their high temporal resolution allows for a direct test of which processing stages---perceptual and/or post-perceptual---are influenced by higher-level capacities such as semantic knowledge. This is made possible by the fact that when participants are presented with visual stimuli, their averaged brain responses show characteristic deflections (ERP components), each of them with its typical polarity (positive or negative), peak latency, spatial scalp distribution, and functional role(s).

The visual P1 component of the ERP refers to a positive deflection peaking as early as 100--150 ms after stimulus onset at occipital channels. It is generated by multiple sources in the extrastriate cortices of the middle occipital and fusiform gyri [@dirusso2001; @mangun1995]. The P1 is thought to reflect the processing of low-level visual characteristics of the stimulus [e.g., size, luminance, and contrast\; @dirusso2001; @luck2014; @johannes1995], which is why it being modulated by higher-level cognitive capacities would challenge a purely bottom-up, encapsulated view of visual perception. It is well-established that the P1 is enhanced when participants pay attention to the spatial location of a stimulus [@luck2000; @mangun1995; @mangun1991] but there is also more recent evidence for higher-level cognitive influences that go beyond spatial attention (see below).

The visual N170 (or N1) component refers to the negative deflection following the P1 and is maximal around 150--200 ms after stimulus onset at occipito-temporal channels. Just as the P1, it is influenced by visual parameters of the stimulus [@johannes1995] and by selective attention, especially when participants are required to discriminate between different visual stimuli [@vogel2000; @mangun1991]. The N170 is enlarged (i.e., more negative) in response to faces [@bentin1996; @rossion2011] and other visual stimuli for which one happens to be an expert [e.g., birds, dogs, or cars\; @gauthier2003a; @tanaka2001]. This category selectivity seems to reflect both the processing of individual visual features of the respective objects (including faces) as well as the holistic configuration of these features [@eimer2011; @jacques2010; @rossion1999; @sagiv2001]. Thus, a modulation of the N170 by cognitive capacities such as knowledge would suggest these capacities having an influence on higher-level, holistic perception of the visual stimulus.

Finally, the N400 component refers to more negative ERP amplitudes when stimuli require more as compared to less resources for semantic processing or integration [@kutas2011; @lau2008; @rabovsky2018]. It begins approximately 300 ms after stimulus onset and is most pronounced at centro-parietal channels. A modulation of the N400 component by cognitive capacities such as knowledge can be taken as evidence for a post-perceptual influence on stimulus processing.

Previous studies have focused on these and other ERP components to investigate which processing stages are indeed influenced by knowledge about visual objects [@abdelrahman2008; @gratton2009; @maier2014; @maier2019; @rossion2002; @rossion2004; @tanaka2001; @samaha2018; @weller2019]. Together, they suggest that not only higher-level ERPs like the semantic N400 component are influenced by acquiring information about objects, but also earlier components that typically reflect bottom-up visual processing. This was the case, for instance, when participants were presented with a range of unfamiliar objects, receiving in-depth verbal descriptions about their function for half of the objects and irrelevant verbal information (i.e., cooking recipes) for the other half of the objects [@abdelrahman2008, Experiment 1]. After this learning phase, the same objects were presented in three different ERP tasks which did not require explicit access to any of the learned semantic information. In all three tasks, ERP amplitudes in response to objects for which in-depth knowledge had been acquired differed from those in response to objects for which this had not been the case. These differences occurred not just in the N400 component, indexing a modulation of post-perceptual semantic processing, but also in the P1 component, suggesting a top-down modulation of lower-level perceptual processing. Interestingly, the ERPs observed for previously unfamiliar objects with newly acquired in-depth knowledge were qualitatively similar to those for untrained but well-known objects. The modulation of the P1 component has recently been replicated when the same objects were presented under circumstances of limited attentional resources in an attentional blink paradigm [@weller2019]. In this study, the (neurophysiological) differences in P1 amplitudes between objects with in-depth versus minimal knowledge were correlated with (behavioral) differences in their detection rate during the attentional blink. This can be taken as further evidence for an influence of semantic knowledge on the early visual perception of unfamiliar objects.

A complementary approach has been taken in a recent study investigating the ERPs in response to familiar (rather than unfamiliar) objects which were rendered difficult to recognize by converting them to two-tone, "Mooney" images [@samaha2018, Experiment 4]. In this experiment, participants were trained with meaningful verbal cues, telling them what kind of objects they should look for in the images, or with a non-meaningful perceptual task, familiarizing them with the images but not with their semantic content. When the EEG was measured in a delayed matching task later on, the two types of training led to differential effects, again suggesting a modulation of lower-level visual perception: Over left posterior electrodes, P1 amplitudes were larger and alpha-band power was higher for meaning-trained images than for perceptually-trained images. As before, this effect of perceiving objects in a semantically informed way seems to be behaviorally relevant, as participants tended respond faster and more accurately to meaning-trained as compared to perceptually-trained images in the matching task. This alternative approach further corroborates an early influence of semantic information on the visual perception of objects.

All the previously mentioned studies have in common that they investigated the effects of semantic knowledge on the visual perception of objects only *after* an extensive learning phase had taken place [@abdelrahman2008; @gauthier2003; @maier2014; @maier2019; @rossion2002; @rossion2004; @samaha2018; @weller2019]. Participants performed one or multiple training sessions during which they encountered each object together with its respective label or description multiple times. The EEG was usually not measured---or at least not analyzed and reported---during these training sessions, which sometimes took place on a separate day [@abdelrahman2008; @maier2014; @maier2019] or were spread out across multiple days or weeks [@rossion2002; @rossion2004]. While experimental designs with such substantive training maximize the chances of detecting even subtle top-down effects of the knowledge acquired, they leave at least three conceptual questions unresolved.

The first question is if we can detect any electrophysiological correlates of *semantic insight*, that is, the critical presentation during which an understanding of the previously unfamiliar object is happening (instead of asking if this understanding leads to differential effects in an orthogonal task later on). The second question is how much learning is actually necessary before reliable top-down effects of knowledge on ERP amplitudes can be obtained: Does it actually take dozens of repetitions per object or may a single exposure to the object together with the relevant semantic information be enough? The third and closely related question is about the nature of the effects of knowledge on perception: Are they reflecting genuine top-down effects of the semantic system altering perception online (while the object is being perceived), or do they merely reflect the (re-)activation of stored representations of the objects which have been altered over the course of the learning process? By measuring the ERPs before, during, and after participants received semantic hints about unfamiliar objects, we were able to provide tentative answers to all three of these questions.

In the two experiments subsequently reported, participants were presented with real-world objects that were presumed to be unfamiliar to most of them. They first viewed each of these objects without any semantic information. This first part served as a naive baseline to rule out that ERPs would differ in response to the objects based solely on low-level visual differences. Next, participants viewed each unfamiliar object for a second time, now preceded by verbal keywords. These keywords could either be matching the typical function of the object, thus making it possible for participants to understand what kind of object they were viewing, or they could be non-matching by describing the function of a different object, thus keeping the perception of the object semantically naive. During this second part, we were able to measure the online influence of semantic insight on object-evoked ERPs as it happened. Finally, the objects were presented for a third time (without keywords, as in the first part) to investigate downstream effects of having acquired semantic knowledge about them---thereby mimicking previous studies [e.g., @abdelrahman2008; @samaha2018; @weller2019]. In all three parts, we examined the influence of semantic information on ERPs associated with lower-level visual perception (P1 component, 100--150 ms), higher-level visual perception (N170 component, 150--200 ms), and semantic processing (N400 component, 400--700 ms).

# Experiment 1

## Method

### Participants

Participants for Experiment 1 were 24 German native speakers (13 female, 11 male) with a mean age of 24 years (range 18 to 31) and no history of psychological disorder or treatment. No a priori power analysis was carried out and the sample size was chosen in line with previous EEG studies on object processing in our lab. All participants were right-handed according to the Edinburgh inventory [@oldfield1971] and reported normal or corrected-to-normal vision. They gave written informed consent before starting the experiment and received a compensation of €8 per hour for participating.

### Materials

Stimuli for Experiments 1 and 2 consisted of 240 grayscale photographs of real-world objects, 120 of which were well-known everyday objects (e.g., a bicycle, a toothbrush), serving as filler stimuli of no interest, whereas the other 120 were rare objects presumed to be unfamiliar to the majority of participants (e.g., a galvanometer, an udu drum). A list of these unfamiliar objects can be found in Appendix A. All stimuli were presented on a light blue background with a size of 207 × 207 pixels on a 19-inch LCD monitor with a resolution of 1,280 × 1,024 pixels and a refresh rate of 75 Hz. At a standardized viewing distance of 90 cm, the images of the objects subtended approximately 3.9 degrees of participants' horizontal and vertical visual angle.

For each unfamiliar object, a pair of keywords---a noun and a verb---was selected, describing the typical function or use of the object in a way that could typically be related to its visual features and their configuration (e.g., voltage, measuring; clay pot, drumming). As our central experimental manipulation, the presentation of half of the objects was preceded by keywords that correctly matched their respective function, whereas the presentation of the other half of the objects was preceded by non-matching keywords belonging to one of the other objects. The matching keywords were expected to induce semantically informed perception (i.e., participants suddenly understanding what kind of object they were viewing), whereas the non-matching keywords were expected to hamper such an understanding and keep the perception of the object semantically naive. All participants saw each unfamiliar object with only one type of keywords (matching or non-matching). This assignment of keywords to objects was counterbalanced across participants so that each object was presented with matching keywords (leading to semantically informed perception) and non-matching keywords (leading to naive perception) to an equal number of participants. The experiment was programmed and displayed using Presentation® software (Neurobehavioral Systems, Inc., Berkeley, CA, www.neurobs.com).

### Procedure

Each experimental session consisted of three parts (see Figure \@ref(fig:exp1-plot)A). In the *pre-insight* part, after written informed consent had been obtained and the EEG had been prepared, all 240 familiar and unfamiliar objects were presented once in random order and without any keywords. Each trial consisted of a fixation cross presented in the middle of the screen for 0.5 s, followed by the presentation of the object until participants made a response or until a time out after 3 s. The inter-trial interval until the presentation of the next fixation cross was 0.5 s and participants took a self-timed break after each block of 60 objects. The task, which was kept the same across all three parts, was to classify each object using one of four response alternatives: (a) "I know what this is or have a strong assumption," (b) "I have an assumption what this is," (c) "I have rather no assumption what this is," or (d) "I don't know what this is and have no assumption." Participants were asked to respond as quickly and as accurately as possible by pressing one out of four buttons with the index or middle finger of their left or right hand, respectively. The mapping of the rating scale to the four buttons (left to right or right to left) was counterbalanced across participants.

```{python def_functions, eval=rerun_preproc}
# Import libraries
import mne
import numpy as np

# Define function for preprocessing a single participant
def preproc(
    vhdr_fname,
    metadata,
    n_components,
    method,
    l_freq,
    h_freq,
    event_id,
    tmin,
    tmax,
    baseline,
    reject,
):
    # Read EEG data in microvolts
    raw = mne.io.read_raw_brainvision(vhdr_fname, scale=1e6, preload=True)
    # Create virtual EOG channels
    raw = mne.set_bipolar_reference(raw, "Auge_u", "Fp1", ch_name="VEOG", drop_refs=False)
    raw = mne.set_bipolar_reference(raw, "F9", "F10", ch_name="HEOG", drop_refs=False)
    raw.set_channel_types(mapping={"VEOG": "eog", "HEOG": "eog"})
    # Add EasyCap electrode layout, removing any excessive channels
    montage = mne.channels.make_standard_montage("easycap-M1")
    raw.drop_channels(list(set(raw.ch_names) - set(montage.ch_names) - set(["VEOG", "HEOG"])))
    raw.set_montage(montage=montage)
    # Re-reference to common average
    raw, _ = mne.set_eeg_reference(raw, "average")
    # Run ICA on a copy of the data
    filt_raw = raw.copy()
    filt_raw.load_data().filter(l_freq=1, h_freq=None)
    ica = mne.preprocessing.ICA(n_components=n_components, random_state=12345, method=method)
    ica.fit(filt_raw)
    # Remove bad components based on correlations with EOG
    eog_indices, eog_scores = ica.find_bads_eog(raw)
    ica.exclude = eog_indices
    raw = ica.apply(raw)
    # Apply band-pass filter
    raw = raw.filter(l_freq=l_freq, h_freq=h_freq)
    # Epoching including baseline correction
    events, _ = mne.events_from_annotations(raw, verbose=False)
    epochs = mne.Epochs(
        raw, events=events, event_id=event_id, tmin=tmin, tmax=tmax, baseline=baseline, preload=True
    )
    # Add behavioral data
    epochs.metadata = metadata
    # Reject bad epochs
    epochs = epochs.drop_bad(reject=reject)
    return epochs

# Define function for preprocessing a list of participants
def preproc_all(
    vhdr_fnames,
    metadatas,
    output_fname,
    n_components,
    method,
    l_freq,
    h_freq,
    event_id,
    tmin,
    tmax,
    baseline,
    reject,
):
    # Do the actual preprocessing
    epochs = [
        preproc(
            vhdr_fname=vhdr_fname,
            metadata=metadata,
            n_components=n_components,
            method=method,
            l_freq=l_freq,
            h_freq=h_freq,
            event_id=event_id,
            tmin=tmin,
            tmax=tmax,
            baseline=baseline,
            reject=reject,
        )
        for vhdr_fname, metadata in zip(vhdr_fnames, metadatas)
    ]
    # Concatenate all participants and save
    epochs = mne.epochs.concatenate_epochs(epochs)
    epochs.save(output_fname, overwrite=True)
    # Return the epochs without loading the actual data
    epochs = mne.read_epochs(output_fname, preload=False)
    return epochs

# Define function to compute averaged waveforms by part and condition
def compute_evokeds(epochs_fname, grand=True, data_only=True):
    # Read preprocessed epochs
    epochs = mne.read_epochs(epochs_fname)
    # Define lists of participants, parts, and conditions
    ps = epochs.metadata["participant"].unique()
    pts = ["I", "II", "III"]
    cns = ["Informed", "Naive"]
    # Define function to compute a single by-participant waveform
    def compute_evoked(epochs, pt, cn, p):
        query = 'participant == "' + p + '" and part == "' + pt + '" and condition == "' + cn + '"'
        return epochs[query].average()

    # Create a dict with by-participant waveforms nested in conditions nested in parts
    avgs = {pt: {cn: [compute_evoked(epochs, pt, cn, p) for p in ps] for cn in cns} for pt in pts}
    # If needed, compute grand-averaged waveforms across participants
    if grand:
        avgs = {pt: {cn: mne.grand_average(avgs[pt][cn]) for cn in cns} for pt in pts}
    # If needed, extract only the data from the evoked MNE objects
    if data_only:
        if grand:
            avgs = {pt: {cn: avgs[pt][cn].data for cn in cns} for pt in pts}
        else:
            ps_num = range(0, len(ps))
            avgs = {pt: {cn: [avgs[pt][cn][p].data for p in ps_num] for cn in cns} for pt in pts}
    return avgs

# Define function to compute single-trial mean ERP amplitudes from epochs
def compute_erps(epochs_fname, name, tmin, tmax, roi):
    # Read preprocessed epochs
    epochs = mne.read_epochs(epochs_fname)
    # Average across electrodes in the ROI
    roi_idx = mne.pick_channels(epochs.ch_names, roi)
    epochs_roi = mne.channels.combine_channels(epochs, dict(roi=roi_idx))
    # Average across samples in the time window
    tmin_s, tmax_s = tmin / 1000, tmax / 1000
    times_idx = np.where(np.logical_and(epochs.times >= tmin_s, epochs.times <= tmax_s))
    erps = epochs_roi.get_data()[:, 0, times_idx].mean(axis=(1, 2))
    return erps

```

>>>>

---
title             : "Instant effects of semantic information on visual perception"
shorttitle        : "Semantically informed perception"
author: 
  - name          : "Alexander Enge"
    affiliation   : "1,2"
    corresponding : yes
    address       : "Rudower Chaussee 18, 12489 Berlin, Germany"
    email         : "alexander.enge@hu-berlin.de"
  - name          : "Franziska Süß"
    affiliation   : "3"
  - name          : "Rasha Abdel Rahman"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "Humboldt-Universität zu Berlin"
  - id            : "2"
    institution   : "Max Planck Institute for Human Cognitive and Brain Sciences"
  - id            : "3"
    institution   : "Fachhochschule des Mittelstands"
authornote: |
  \addORCIDlink{Alexander Enge}{0000-0003-0100-2297}

@@

  The authors declare no competing financial interests.
  The preprocessed data and analaysis code for this study are openly available at https://osf.io/uksbc/.
abstract: |
  Does our perception of an object change as soon as we discover what function it serves?
  This question matters not only when we encounter novel tools or gadgets in our everyday lives, but also pertains to the long-standing debate around the (im)penetrability of perception by higher cognitive capacities.
  Here, we showed human participants (*n* = 48) pictures of unfamiliar objects either together with words matching their function, leading to semantically informed perception, or together with non-matching words, resulting in naive perception.
  We measured event-related potentials (ERPs) to investigate at which stages in the visual processing hierarchy these two types of object perception differed from one another.
  We found that semantically informed as compared to naive perception was associated with larger amplitudes in the N170 component (150--200 ms) and reduced amplitudes in the N400 component (400--700 ms).
  When the same objects were presented once more without any information, the N400 effect persisted and we also observed enlarged amplitudes in the P1 component (100--150 ms) in response to objects for which semantically informed perception had taken place.
  Consistent with previous work, our results suggest that obtaining semantic information about previously unfamiliar objects alters aspects of their lower-level visual perception (P1 component), higher-level visual perception (N170 component), and semantic processing (N400 component).
  We demonstrate for the first time that these effects are instantaneous in the sense that they can be observed within the same trial as well as one trial after the information has been given, providing strong evidence that cognition affects perception.
keywords          : "objects, semantic knowledge, visual perception, event-related potentials"
floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
bibliography      : ["references.bib", "r-references.bib"]
csl               : "apa.csl"
documentclass     : "apa7"
classoption       : "man,donotrepeattitle"
fontsize          : "10pt"
output:
  papaja::apa6_pdf:
    latex_engine: xelatex
  papaja::apa6_docx: default
editor_options: 
  chunk_output_type: console
header-includes:
  - \geometry{a4paper}
  - \fancyheadoffset[R,L]{0pt}
  - \raggedbottom
  - \usepackage[all]{nowidow}
  - \usepackage[bottom]{footmisc}
  - \interfootnotelinepenalty=10000
  - \usepackage{setspace}
  - \AtBeginEnvironment{tabular}{\doublespacing}
  - \usepackage{makecell}
  - \renewcommand{\cellset}{\renewcommand{\arraystretch}{0.7}}
  - \captionsetup{font={stretch=1}, belowskip=15pt}
---

```{r, setup, include=FALSE}
# Load R packages
library(buildmer)
library(emmeans)
library(here)
library(knitr)
library(lme4)
library(lmerTest)
library(MASS)
library(papaja)
library(parallel)
library(reticulate)
library(tidyverse)
library(magrittr)

# Set global chunk options
opts_chunk$set(fig.width = 10, include = FALSE, out.width = "100%")

# Write R packages to bibliography
r_refs(here("analysis", "r-references.bib"), append = FALSE)

# What to (re-)run
run <- list(
  "preprocessing" = TRUE,
  "modeling" = TRUE
)

# Directory paths
data_dir <- here("data")
output_dir <- here("output")
```

```{r, preprocessing, eval=run$preprocessing}
# Process behavioral data
c(
  list.files(here(data_dir, "rt/exp1"), "*.txt", full.names = TRUE),
  list.files(here(data_dir, "rt/exp2"), "*.txt", full.names = TRUE)
) %>%
  # Process each log file
  map(function(log_file) {

    # Read the file
    read_tsv(
      log_file,
      col_types = cols(),
      locale = locale(encoding = "latin1")
    ) %>%
      # Recode some columns
      transmute(

        # Phase of the experiment
        phase = factor(Wdh,
          levels = c(211, 212, 213),
          labels = c("Pre-insight", "Insight", "Post-insight")
        ),

        # # Participant ID (unique for Experiments 1 and 2)
        # participant = ifelse(
        #   str_detect(log_file, "exp2"), str_c(VPNummer, "_exp2"), VPNummer
        # ),

        # Keyword conditions
        keywords = factor(Bed,
          levels = c("richtig", "falsch"),
          labels = c("Match", "Non-match")
        ),

        # Behavioral responses
        # 201: "I know what this is or have a strong assumption"  -> Recode as 4
        # 202: "I have an assumption what this is"                -> Recode as 3
        # 203: "I have rather no assumption what this is"         -> Recode as 2
        # 204: "I don't know what this is and have no assumption" -> Recode as 1
        response = 5 - (Tastencode - 200),

        # Reaction times
        rt = RT,
        log_rt = log(RT),

        # Item IDs
        item_id = factor(StimID)
      ) %>%
      # Remove filler items
      drop_na(keywords) -> log

    # Assign items to conditions based on manipulation (matching vs. non-
    # matching keywords) and the responses of the participants
    items_per_condition <- with(log, {
      list(

        # Informed condition: Matching keywords and positive response
        "Informed" =
          item_id[phase == "Insight" & keywords == "Match" & response >= 3],

        # Naive condition: Non-matching keywords and negative response
        "Naive" =
          item_id[phase == "Insight" & keywords == "Non-match" & response <= 2],

        # Exclude: Matching keywords and negative response
        "Exclude_informed" =
          item_id[phase == "Insight" & keywords == "Match" & response <= 2],

        # Exclude: Non-matching keywords and positive response
        "Exclude_naive" =
          item_id[phase == "Insight" & keywords == "Non-match" & response >= 3],

        # Exclude: Positive response *before* any info was presented
        "Exclude_known" = item_id[phase == "Pre-insight" & response == 4]
      )
    })

    # Assign these conditions to the trials of all three phases
    log$condition <- NA
    for (condition in names(items_per_condition)) {
      item_id <- items_per_condition[[condition]]
      log$condition[log$item_id %in% item_id] <- condition
    }

    # Sort columns
    log %>%
      select(item_id, phase, condition, response, rt, log_rt) %>%
      return()
  }) -> logs

# Install latest version of the EEG pipeline
system("pip install --user --upgrade --pre --index https://test.pypi.org/simple/ hu-neuro-pipeline")

# Import EEG pipeline from Python
pipeline <- import("pipeline")

# Get list of BrainVision EEG header files
vhdr_files <- c(
  list.files(here(data_dir, "eeg/exp1"), "*.vhdr", full.names = TRUE),
  list.files(here(data_dir, "eeg/exp2"), "*.vhdr", full.names = TRUE)
)

# Process EEG data
res <- pipeline$group_pipeline(
  vhdr_files = vhdr_files,
  log_files = logs,
  export_dir = output_dir,
  downsample_sfreq = 250.,
  # bad_channels = "auto",
  ocular_correction = "fastica",
  highpass_freq = 0.1,
  lowpass_freq = 40.,
  triggers = c(221, 222),
  epochs_tmin = -0.5,
  epochs_tmax = 1.5,
  baseline = c(-0.2, 0.0),
  reject_peak_to_peak = 200.,
  reject_flat = 1.,
  components = list(
    "name" = c("P1", "N170", "N400"),
    "tmin" = c(0.1, 0.15, 0.4),
    "tmax" = c(0.15, 0.2, 0.7),
    "roi" = list(
      c("PO3", "PO4", "POz", "O1", "O2", "Oz"),
      c("P7", "P8", "PO7", "PO8", "PO9", "PO10"),
      c("C1", "C2", "Cz", "CP1", "CP2", "CPz")
    )
  ),
  average_by = c("phase/condition"),
  perform_tfr = TRUE,
  perm_contrasts = list(
    c("Pre-insight/Informed", "Pre-insight/Naive"),
    c("Insight/Informed", "Insight/Naive"),
    c("Post-insight/Informed", "Post-insight/Naive")
  ),
  n_jobs = 1
)

# Extract results
trials <- tibble(res[[1]])
evokeds <- tibble(res[[2]])
clusters <- tibble(res[[3]])
tfr_evokeds <- tibble(res[[4]])
tfr_clusters <- tibble(res[[5]])
```

```{r, modeling, eval=run$modeling}
# Read single trial data
read_csv(here(output_dir, "all_trials.csv"), col_types = cols(
  participant_id = col_factor(),
  item_id = col_factor(),
  phase = col_factor(),
  condition = col_factor()
)) %>%
  # Subset only the relevant conditions
  mutate(condition = factor(
    condition,
    levels = c("Informed", "Naive", "Excl_informed")
  )) %>%
  drop_na() -> trials

# Contrast coding for phase
cbind(
  c("Pre-insight" = -1, "Insight" = 1, "Post-insight" = 0),
  c("Pre-insight" = 0, "Insight" = -1, "Post-insight" = 1)
) %>%
  t() %>%
  ginv() -> contrasts(trials$phase)

# Contrast coding for condition
cbind(
  c("Informed" = 1, "Naive" = -1, "Excl_informed" = 0),
  c("Informed" = 1, "Naive" = 0, "Excl_informed" = -1)
) %>%
  t() %>%
  ginv() -> contrasts(trials$condition)

# Prepare model formula
mod_formula <- N170 ~ phase * condition +
  (phase * condition | participant_id) +
  (phase * condition | item_id)

# Fit linear mixed models
deps <- c("P1", "N170", "N400")
mod_lists <- map(deps, function(dep, mod_formula = form, mod_data = trials) {

  # Find the best fitting model (see Matuschek et al., 2017, *JML*)
  # Uses parallelization across all available CPU cores
  n_cores <- detectCores()
  cl <- makeCluster(n_cores)
  build <- buildmer(
    buildmerControl = buildmerControl(
      formula = mod_formula,
      data = mod_data,
      # args = list("control" = lmerControl(
      #   optimizer = "bobyqa", optCtrl = list(maxfun = 1e6)
      # )),
      direction = c("order", "backward"),
      cl = cl,
      elim = LRTalpha(.20),
      calc.anova = TRUE,
      ddf = "Satterthwaite",
      dep = dep
    )
  )
  stopCluster(cl)

  # Compute marginal means and contrasts
  emm_options(lmer.df = "Satterthwaite", lmerTest.limit = Inf)
  specs <- pairwise ~ condition | phase
  em <- emmeans(build@model, specs, infer = TRUE)

  # Create list of returns
  list(
    "model" = build@model,
    "summary" = build@summary,
    "anova" = build@anova,
    "means" = em$emmeans,
    "contrasts" = em$contrasts
  ) %>%
    return()
}) %>%
  # Add ERP component names
  set_names(deps)

# Save models
saveRDS(mod_lists, here(output_dir, "models.RDS"))
```
